---
title: "Assignment1"
author: "Clarisa Griffin"
date: "2025-01-08"
output: html_document
---

[GitHub Repository: Assignment 1](https://github.com/CEGriffin/BIOL432_Assignment1)

This script generates limb length and width data of 5 species of Patagonian caracaras, then exports the data to a measurements.csv
```{r}
source("dataGenerato.R", local=knitr::knit_global(), echo=T)
```

This script uses the limb length and width data generated in the previous script to generate limb volume data of 5 species of Patagonian caracaras, then exports the data to measurements.csv
```{r}
source("volumeEstimato.R", local=knitr::knit_global(), echo=T)
```

Load libraries
```{r}
library(dplyr)
library(ggplot2)
source("http://bit.ly/theme_pub")
library(tidyverse)
```


Load measurements.csv
```{r}
limb_data<-read.csv("measurements.csv", header=T)
```


Sort data by species, then observer, then limb volume
```{r}
limb_data<-limb_data %>%
  arrange(spec_names, observer, limb_volume_cm3)
```


Table showing average limb volume for each species
```{r}
#create table
limb_table<-limb_data %>%
  group_by(spec_names) %>%
  summarize("Average limb volume (cm^3)"=round(mean(limb_volume_cm3),2))

#rename columns
colnames(limb_table)[colnames(limb_table)=="spec_names"]="Species names"

print(limb_table)
```


Table showing number of observations for each combination of species and observer
```{r}
obs_table<-table(limb_data$spec_names, limb_data$observer)

#rename columns and rows
colnames(obs_table)=c("B. Sanderson", "N. Gaiman", "V. E. Schwab")
rownames(obs_table)=c("Chimango", "Crested", "Mountain", "Northern","Striated")

print(obs_table)
```


Boxplot comparing mean, standard deviations, and maximum and minimums of the distributions of volumes for each species

Create a custom function to calculate the summary statistics of each distribution
```{r}
sum_stats<-function(x){
  output<-c(min(x),
            mean(x, na.rm=T)-sd(x, na.rm=T)/(sqrt(length(x))),
            mean(x, na.rm=T),
            mean(x, na.rm=T)+sd(x, na.rm=T)/(sqrt(length(x))),
            max(x)
  )
  names(output)<-c("ymin", "lower", "middle", "upper", "ymax")
  output
}
```


Show limb volume distributions in a boxplot
```{r}
limb_vol_plot<-ggplot(data=limb_data, aes(x=spec_names, y=limb_volume_cm3)) +
  theme_pub() +
  stat_summary(fun.data=sum_stats, geom="boxplot") +
  labs(x="Species names", y="Limb volume (cm^3)") +
  scale_x_discrete(labels=c("Chimango","Crested","Mountain","Northern","Striated"))
print(limb_vol_plot)
```



Multipanel frequency histogram plot showing distribution of limb volume for each species

Create vectors to rename the species within the plot
```{r}
caracara_names<-c("Chimango","Crested","Mountain","Northern","Striated")
names(caracara_names)<-c("chimango_caracara", "crested_caracara", "mountain_caracara", "northern_caracara", "striated_caracara")
```


Create frequency plot
```{r}
limb_freq_plot<-ggplot(data=limb_data, aes(x=limb_volume_cm3)) +
  theme_pub() +
  geom_histogram(bins=10) +
  facet_grid(~spec_names, scales="free", labeller=labeller(spec_names=caracara_names)) +
  labs(x="Limb volume (cm^3)", y="Count")
print(limb_freq_plot)
```

